---
title: "Figure 6.1 - Label transfer workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{condor_workflow}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup}
library(condor)
library(ggpubr)
library(ggrastr)
```

```{r}
cluster_palette <- c("#89C5DA", "#DA5724", "#74D944", "#CE50CA", "#3F4921", "#C0717C", "#CBD588", "#5F7FC7",
                     "#673770", "#D3D93E", "#38333E", "#508578", "#D7C1B1", "#689030", "#AD6F3B", "#CD9BCD",
                     "#D14285", "#6DDE88", "#652926", "#7FDCC0", "#C84248", "#8569D5", "#5E738F", "#D1A33D",
                     "#8A7C64", "#599861", "#89C5DA", "#DA5724", "#74D944", "#CE50CA", "#3F4921", "#C0717C", "#CBD588", "#5F7FC7",
                     "#673770", "#D3D93E", "#38333E", "#508578", "#D7C1B1", "#689030", "#AD6F3B", "#CD9BCD",
                     "#D14285", "#6DDE88", "#652926", "#7FDCC0", "#C84248", "#8569D5", "#5E738F", "#D1A33D",
                     "#8A7C64", "#599861")
```

# Loading the data for training
```{r}
condor_train <- prep_fcd(FCSpath = "/home/user/data/Data/Package/condor/.test_files/fcs/", 
                        ceil = 5000, 
                        useCSV = FALSE, 
                        transformation = "a", 
                        remove_param = c("FSC-H", "SSC-H", "FSC-W", "SSC-W", "Time", "InFile", "live_dead"), 
                        anno_table = "./data_and_envs/metadata_train.csv", 
                        filename_col = "filename",
                        seed = 91)
```

```{r}
class(condor_train)
```

```{r}
condor_train$anno$cell_anno$group <- "train"
```

# Loading the data
```{r}
condor_test <- prep_fcd(FCSpath = "/home/user/data/Data/Package/condor/.test_files/fcs/", 
                        ceil = 5000, 
                        useCSV = FALSE, 
                        transformation = "a", 
                        remove_param = c("FSC-H", "SSC-H", "FSC-W", "SSC-W", "Time", "InFile", "live_dead"), 
                        anno_table = "./data_and_envs/metadata_test.csv", 
                        filename_col = "filename",
                        seed = 91)
```

```{r}
class(condor_test)
```

```{r}
condor_test$anno$cell_anno$group <- "test"
```

# UMAP Projection

## Run UMAP keeping the model
```{r}
condor_train <- runUMAP(fcd = condor_train, 
                        input_type = "expr", 
                        data_slot = "orig", 
                        seed = 91, 
                        n_threads = 4, 
                        ret_model = TRUE)
```

## Add data to the embedding
```{r}
condor_test <- learnUMAP(fcd = condor_test, 
                         input_type = "expr", 
                         data_slot = "orig", 
                         model = condor_train$extras$umap_model, 
                         n_epochs = 100, 
                         seed = 91, 
                         n_threads = 4)
```

# Train a classifier for the label transfer
```{r}
condor_train <- runPhenograph(fcd = condor_train, 
                              input_type = "expr", 
                              data_slot = "orig", 
                              k = 300, 
                              seed = 91)
```

## Visualization
```{r}
plot_marker(data = cbind(condor_train$umap$expr_orig, condor_train$clustering$Phenograph_expr_orig_k_300), 
            param = "Phenograph", 
            order = T, 
            title = "Figure S9b -UMAP on the training dataset, Phenograph clusters", 
            dim_red = "UMAP", 
            facet_by_variable = FALSE, 
            label_clusters = FALSE)
```

### Rastered Phenograph clustering
```{r}
data <- cbind(condor_train$umap$expr_orig, condor_train$anno$cell_anno, condor_train$clustering$Phenograph_expr_orig_k_300)
```

```{r}
ggplot(data = data, aes(x = UMAP1, y = UMAP2, color = Phenograph)) +
  geom_point_rast(alpha = 0.2, size = 0.1) +
  theme_bw() +
  theme(aspect.ratio = 1, panel.grid = element_blank()) +
  scale_color_manual(values = cluster_palette) + 
  guides(color = guide_legend(override.aes = list(size = 5, alpha = 1))) + 
  ggtitle("Figure S9b -UMAP on the training dataset, Phenograph clusters")
```

# Metaclustering
```{r}
condor_train <- metaclustering(fcd = condor_train, 
                               clustering = "Phenograph_expr_orig_k_300", 
                               name_col = "Phenograph", 
                               name_out = "metaclusters", 
                               metaclusters = c("1" = "Classical Monocytes", 
                                                "2" = "CD8 T", 
                                                "3" = "CD4 T", 
                                                "4" = "Classical Monocytes", 
                                                "5" = "NK", 
                                                "6" = "CD8 T", 
                                                "7" = "CD4 T", 
                                                "8" = "CD4 T", 
                                                "9" = "Non-classical Monocytes", 
                                                "10" = "CD8 T", 
                                                "11" = "Classical Monocytes", 
                                                "12" = "B", 
                                                "13" = "CD4 T",
                                                "14" = "pDC"))
```

```{r}
plot_marker(data = cbind(condor_train$umap$expr_orig, condor_train$clustering$Phenograph_expr_orig_k_300), 
            param = "metaclusters", 
            order = T, 
            title = "Figure 6b - UMAP on the training dataset, metaclusters", 
            dim_red = "UMAP", 
            facet_by_variable = FALSE, 
            label_clusters = FALSE)
```

### Rastered metaclustering
```{r}
data <- cbind(condor_train$umap$expr_orig, condor_train$anno$cell_anno, condor_train$clustering$Phenograph_expr_orig_k_300)
```

```{r}
ggplot(data = data, aes(x = UMAP1, y = UMAP2, color = metaclusters)) +
  geom_point_rast(alpha = 0.2, size = 0.1) +
  theme_bw() +
  theme(aspect.ratio = 1, panel.grid = element_blank()) +
  scale_color_manual(values = cluster_palette) + 
  guides(color = guide_legend(override.aes = list(size = 5, alpha = 1))) + 
  ggtitle("Figure 6b - UMAP on the training dataset, metaclusters")
```

# Label transfer

## Train label trasfer kNN clussifier
```{r}
condor_train <- train_transfer_model(fcd = condor_train, 
                                     data_slot = "orig", 
                                     input_type = "expr", 
                                     label = condor_train$clustering$Phenograph_expr_orig_k_300$metaclusters, 
                                     method = "knn", 
                                     tuneLength = 5, 
                                     trControl = caret::trainControl(method = "cv"), 
                                     seed = 91)
```

```{r}
condor_train$extras$lt_model$performace_plot + ggtitle("Figure S11a - kNN accuracy")
```

```{r, fig.height=10, fig.width=10}
# Figure S11b - kNN importance

condor_train$extras$lt_model$features_plot
```

## Trasfer lables on the new data
```{r}
condor_test <- predict_labels(fcd = condor_test, 
                              data_slot = "orig", 
                              input_type = "expr", 
                              model_object = condor_train, 
                              label = "label_pred", 
                              seed = 91)
```

## Train label transfer kNN classifier - cluster level
```{r}
condor_train_fine <- train_transfer_model(fcd = condor_train, 
                                          data_slot = "orig", 
                                          input_type = "expr", 
                                          label = condor_train$clustering$Phenograph_expr_orig_k_300$Phenograph, 
                                          method = "knn", 
                                          tuneLength = 5, 
                                          trControl = caret::trainControl(method = "cv"), 
                                          seed = 91)
```

```{r}
condor_train_fine$extras$lt_model$performace_plot + ggtitle("Figure S10a - kNN accuracy")
```

```{r, fig.height=10, fig.width=10}
# Figure S10b - kNN importance

condor_train_fine$extras$lt_model$features_plot
```

```{r}
condor_test <- predict_labels(fcd = condor_test, 
                              data_slot = "orig", 
                              input_type = "expr", 
                              model_object = condor_train_fine, 
                              label = "label_pred_fine", 
                              seed = 91)
```

# Manual annotation of the dataset
```{r}
condor_test <- runPhenograph(fcd = condor_test, 
                             input_type = "expr", 
                             data_slot = "orig", 
                             k = 10, 
                             seed = 91)
```

## Metaclustering
```{r}
condor_test <- metaclustering(fcd = condor_test, 
                              clustering = "Phenograph_expr_orig_k_10", 
                              name_col = "Phenograph", 
                              name_out = "metaclusters", 
                              metaclusters = c("1" = "CD8 T", 
                                               "2" = "CD8 T", 
                                               "3" = "Classical Monocytes", 
                                               "4" = "CD4 T", 
                                               "5" = "Classical Monocytes", 
                                               "6" = "CD8 T", 
                                               "7" = "NK", 
                                               "8" = "Classical Monocytes", 
                                               "9" = "Classical Monocytes", 
                                               "10" = "CD4 T", 
                                               "11" = "NK", 
                                               "12" = "CD4 T", 
                                               "13" = "CD4 T",
                                               "14" = "CD4 T",
                                               "15" = "CD8 T",
                                               "16" = "Classical Monocytes",
                                               "17" = "Non-classical Monocytes",
                                               "18" = "CD8 T",
                                               "19" = "B",
                                               "20" = "CD8 T",
                                               "21" = "pDC"))
```

```{r, fig.width=10, fig.height=10}
confusion_HM(variables = condor_test$clustering$Phenograph_expr_orig_k_10$metaclusters, 
             group = condor_test$clustering$label_pred$predicted_label, size = 30, title = "Figure 6f - confusion matrix")
```

# Plotting for figure 6

## Prepare the dataframe
```{r}
train <- cbind(condor_train$umap$expr_orig, condor_train$clustering$Phenograph_expr_orig_k_300[, c(1,3)])

train$type <- "train"

test <- cbind(condor_test$umap$expr_orig, 
              condor_test$clustering$label_pred_fine,
              condor_test$clustering$label_pred)

test$Description <- NULL
test$Description <- NULL

colnames(test) <- c("UMAP1", "UMAP2", "Phenograph", "metaclusters")

test$type <- "test"

vis_data <- rbind(train, test)
```

# Overlap UMAP
```{r}
vis_data$type <- factor(vis_data$type, levels = c("train", "test"))
```

```{r}
ggplot(data = vis_data, aes(x = UMAP1, y = UMAP2, color = type, alpha = type, size = type)) +
  geom_point_rast() +
  scale_color_manual(values = c("gray", "#92278F")) +
  scale_alpha_manual(values = c(0.5, 1)) +
  scale_size_manual(values = c(0.1, 0.5)) +
  theme_bw() +
  theme(aspect.ratio = 1, panel.grid = element_blank()) + 
  ggtitle("Figure 6c - UMAP projected")
```

```{r}
ggplot(data = vis_data, aes(x = UMAP1, y = UMAP2, color = metaclusters, alpha = type, size = type)) +
  geom_point_rast() +
  scale_color_manual(values = cluster_palette) +
  scale_alpha_manual(values = c(0.01, 1)) +
  scale_size_manual(values = c(0.01, 0.1)) +
  theme_bw() +
  theme(aspect.ratio = 1, panel.grid = element_blank()) + 
  ggtitle("Figure 6e - Predicted cell labels") + facet_wrap(~type)
```

```{r}
ggplot(data = vis_data, aes(x = UMAP1, y = UMAP2, color = Phenograph, alpha = type, size = type)) +
  geom_point_rast() +
  scale_color_manual(values = cluster_palette) +
  scale_alpha_manual(values = c(0.01, 1)) +
  scale_size_manual(values = c(0.01, 0.1)) +
  theme_bw() +
  theme(aspect.ratio = 1, panel.grid = element_blank()) + 
  ggtitle("Figure S9c - Predicted clusters") + facet_wrap(~type)
```

# Calculate LISI score
```{r}
# devtools::install_github("immunogenomics/lisi")

library(lisi)
```

## Prepare dataset
```{r}
train_filt <- train[sample(rownames(train), 5000, replace = FALSE),]

lisi_mat <- rbind(train_filt, test)

res <- compute_lisi(lisi_mat[,c(1,2)], lisi_mat, c('type'))

colnames(res) <- "lisi"

lisi_mat <- cbind(lisi_mat, res)
```

# Visualization
```{r}
ggplot(data = lisi_mat, aes(y = lisi, x = "LISI")) + 
  geom_jitter_rast(alpha = 0.3) +
  geom_violin(alpha = 0.8) +
  theme_bw() +
  theme(aspect.ratio = 2, panel.grid = element_blank()) + ggtitle("Figure 6d - LISI global")
```

```{r}
ggplot(data = lisi_mat, aes(y = lisi, x = metaclusters, fill = metaclusters)) + 
  geom_jitter_rast(alpha = 0.3) +
  geom_violin() +
  theme_bw() +
  scale_fill_manual(values = cluster_palette) +
  theme(aspect.ratio = 1/4, panel.grid = element_blank())+ ggtitle("Figure 6d - LISI metaclusters")
```

```{r}
ggplot(data = lisi_mat, aes(y = lisi, x = Phenograph, fill = Phenograph)) + 
  geom_jitter_rast(alpha = 0.3) +
  geom_violin() +
  theme_bw() +
  scale_fill_manual(values = cluster_palette) +
  theme(aspect.ratio = 1/4, panel.grid = element_blank())+ ggtitle("Figure 6d - LISI clusters")
```


# Session Info
```{r}
info <- sessionInfo()

info
```
